---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# exirt

<!-- badges: start -->
[![R-CMD-check](https://github.com/UO-BRT/exirt/workflows/R-CMD-check/badge.svg)](https://github.com/UO-BRT/exirt/actions)
<!-- badges: end -->

The **exirt** is designed to automate annual analyses for the Oregon Extended 
Assessment. It is the analysis arm of a suite of packages designed to (a)
access the live data via a web API ([orextdb](https://github.com/UO-BRT/orextdb)),
(b) process the data for different analyes and visualizaionts 
([dbprocess](https://github.com/UO-BRT/dbprocess)), and (c) analyze the data
([exirt](https://github.com/UO-BRT/exirt)).

Here are the three packages created by Daniel Anderson and maintained by Christopher Loan:

    * `{orextdb}` — accesses data base
    * `{dbprocess}` — processes data
    * `{exirt}` — fit irt models, plot data, etc.

# Getting Set Up

These are stored on github in the UO-BRT organization. To install these, use `devtools::install_github()` with the organization and package name, separated by a slash:

## Install Packages

```{r eval = FALSE}
devtools::install_github('UO-BRT/orextdb')
devtools::install_github('UO-BRT/dbprocess')
devtools::install_github('UO-BRT/exirt')
```

## Load these packages

You'll see I'm also loading the `{tidyverse}` for data manipulation.

```{r message = FALSE}
library(orextdb)
library(dbprocess)
library(exirt)
library(tidyverse)
```

# `{orextdb}`

## Accessing the Data Base

You can check if you have access with `db_key()`. If this returns an empty string, you do not have the key. You can set the key by wrapping the key in quotes within `db_set_key()` (e.g., `db_set_key('asdfjkl-123456')`)

Let's work with 2018-19 for our purposes. databases are specified with the prefix "ORExt" and then the final two digits of the academic year. So for 2018-19, we specify `db = 'ORExt1819'`. 


We also have to specify which database we want. If we want to estimate the ability estimates, we need to specify which table we want pulled from the database. 

This is done with the argument `table = `, which is a string specifying the specific table from the Oregon Extended live database. 

`table` should be one of 

    * "Accomodations"
    * "Answers"
    * "Districts" 
    * "Exams"
    * "Items" 
    * "Preferences" 
    * "Schools"
    * "Students" 
    * "Students_old" 
    * "Submissions" 
    * "SupplementalDistricts"
    * "SupplementalSchools"
    * "Tasks"
    * "User" 
    * "UserStudents"
    * "UserStudents_old"


Check the documentation with `?db_get()` for further arguments, though defaults have been set to be useful and will not need to be changed typically.


```{r}
answers_1819 <- 
  orextdb::db_get(
    table = 'Answers',
    db = 'ORExt1819'
    )

answers_1819
```

```{r}
students_1819 <- 
  orextdb::db_get(
    table = 'Students',
    db = 'ORExt1819'
    )
```

# `{dbprocess}`

The main function in `{dbprocess}` is `get_items()`. This function pulls item level data with student demographics from the live ORExt database. You have the option to specify a grade, content area, and if you want demographics or not. If you do not specify, all grade and content areas are selected and demographics are included.

Be sure to specify your database, if you don't want it to pull based on your date (which pulls the current year).

```{r}
math_items <- 
  dbprocess::get_items(
    content = 'Math', 
    grade = 11,
    db = 'ORExt1819'
    )

```

# `{exirt}`

## estimate abilities (omitting field test)

We use `estimate_abilities()` provide test score data with `test` to determine the ability estimates for all non-field-test items. We'll continue with only grade 11 math, but a list of test score data can be passed to this argument.

```{r}
math_abilities <- 
  exirt::estimate_abilities(
    test = math_items
    )
```

## estimate field test difficulty

With very similar specification to `estimate_abilities()`, we can `estimate_ft_difficulties()` to determine the difficulty parameters of the field-test items

```{r}
math_ft_difficulty <- 
  exirt::estimate_ft_difficulties(
  test = math_items
)
```

## other functions

### fit Rasch models

The `rasch()` function his is mostly an internal function, called by the functions above, but may be useful to someone wanting additional model information. It uses `TAM::tam.mml()` to run the Rasch models. You can tell this to omit the field test items or anchor the other items.

```{r}
rasch_mod <- 
  exirt::rasch(
    test = math_items, 
    omit_field_test = TRUE
  )
```

### pulling information from Rasch model

We can get all of the field test item names from the Rasch model with the function `get_ft_items()`

```{r}
ft_item_names <- exirt::get_ft_items(rasch_mod)
```
We can then re-fit the rasch model without omitting the field test items. 

```{r}
rasch_mod_ft <- 
  exirt::rasch(
    test = math_items, 
    omit_field_test = FALSE
  )
```

If we use the names of the field test items (defined above), we can then fitler out the item difficulties, so we only look at the field test items.

```{r}
exirt::get_item_diffs(rasch_mod_ft) %>% 
  filter(item_id %in% ft_item_names$item_id)
```


It might be useful to have a list of all raw scores, thetas, RIT scores, and corresponding performance levels. All of this can be returned with `exirt::raw_to_rit()`. This returns for all tests and grades, and can be filtered by whatever grade you want with `df %>% filter(test %in% c('Math_G3', 'ELA_G3'))` for example.

```{r}
raw_to_rit_scores <- 
  exirt::raw_to_rit()

raw_to_rit_scores %>% 
  filter(test %in% c('Math_G3', 'ELA_G3')) %>% 
  head()
```
If you want these for the ELA subscores, you can repeat the call, but add the `subscore = TRUE` argument (the default is `FALSE`, giving you the Math, ELA, and Science scores).

```{r}
raw_to_rit_subscores <- 
  exirt::raw_to_rit(subscore = TRUE)

raw_to_rit_subscores %>% 
  head()
```
### Item difficulties

```{r}
item_difficulties <- 
  exirt::get_item_diffs(model_ob = rasch_mod)

item_difficulties %>% 
  head()
```

### Person estimates

Note, I'm not printing these for privacy

```{r}
person_estimates <- 
  get_person_estimates(
    model_ob = rasch_mod,
    full_demo_data = math_items
    )
```


## Plotting 

### Test characteristic curves

```{r message=FALSE, warning=FALSE}
tcc_plot <- 
  exirt::tcc_plot(
    item_diff_table = item_difficulties, 
    content = 'Math', 
    grades = "11"
    )

tcc_plot
```

### Test information functions

```{r message=FALSE, warning=FALSE}
tif_plot <- 
  exirt::tif_plot(item_diff_table = item_difficulties) 

tif_plot
```

